\propertitle

### 2019-06-13 maxNt_v5.tex
update

\pdfoutput=1
%% Author: PGL  Porta Mana
%% Created: 2015-05-01T20:53:34+0200
%% Last-Updated: 2019-06-13T15:02:32+0200
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newif\ifarxiv
\arxivfalse
\ifarxiv\pdfmapfile{+classico.map}\fi
\newif\ifafour
\afourfalse% true = A4, false = A5
\newif\iftypodisclaim % typographical disclaim on the side
\typodisclaimtrue
\newcommand*{\memfontfamily}{zplx}
\newcommand*{\memfontpack}{newpxtext}
\documentclass[\ifafour a4paper,12pt,\else a5paper,10pt,\fi%extrafontsizes,%
onecolumn,oneside,article,%french,italian,german,swedish,latin,
british%
]{memoir}
\newcommand*{\updated}{\today}
\newcommand*{\firstdraft}{4 November 2015}
\newcommand*{\firstpublished}{***}
\newcommand*{\propertitle}{Maximum-entropy distributions for a 
  neuronal population\\ from subpopulation data%\\{\large ***}%
}
\newcommand*{\pdftitle}{\propertitle}
\newcommand*{\headtitle}{Maximum-entropy from sample data}
\newcommand*{\pdfauthor}{P.G.L.  Porta Mana, Y. Roudi, V. Rostami, E. Torre}
\newcommand*{\headauthor}{Porta Mana \etal}
\newcommand*{\reporthead}{}% Report number

@@

This step is even more straightforward if instead of power moments we use
\emph{normalized factorial moments}. The $m$th normalized factorial moment
is defined as the expectation
\begin{equation}
  \E\condt[\Bigg]{
    \binom{n \ya}{m}\,\binom{n}{m}^{-1}}{\yH}
  \equiv
  \sum_{\ya} %m!\,
  \binom{n \ya}{m}\,  \binom{n}{m}^{-1}\, \pf(\ya \| \yH),
\end{equation}
where $\ya\in\set{0, 1/n, 2/n, \dotsc, (n-1)/n, 1}$ is the normalized total
activity of a population of $n$ neurons. This moment can be interpreted as
the expectation of the number of distinct $m$-tuples of simultaneously
active neurons, normalized by the maximum possible number of distinct
$m$-tuples.

>>>>

\pdfoutput=1
%% Author: PGL  Porta Mana
%% Created: 2015-05-01T20:53:34+0200
%% Last-Updated: 2019-06-13T16:07:01+0200
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newif\ifarxiv
\arxivfalse
\ifarxiv\pdfmapfile{+classico.map}\fi
\newif\ifafour
\afourfalse% true = A4, false = A5
\newif\iftypodisclaim % typographical disclaim on the side
\typodisclaimtrue
\newcommand*{\memfontfamily}{zplx}
\newcommand*{\memfontpack}{newpxtext}
\documentclass[\ifafour a4paper,12pt,\else a5paper,10pt,\fi%extrafontsizes,%
onecolumn,oneside,article,%french,italian,german,swedish,latin,
british%
]{memoir}
\newcommand*{\updated}{\today}
\newcommand*{\firstdraft}{4 November 2015}
\newcommand*{\firstpublished}{***}
\newcommand*{\propertitle}{Maximum-entropy distributions for a 
  neuronal population\\ from subpopulation data%\\{\large ***}%
}
\newcommand*{\pdftitle}{\propertitle}
\newcommand*{\headtitle}{Maximum-entropy from sample data}
\newcommand*{\pdfauthor}{P.G.L.  Porta Mana, Y. Roudi, V. Rostami, E. Torre}
\newcommand*{\headauthor}{Porta Mana \etal}
\newcommand*{\reporthead}{}% Report number

@@

This step is even more straightforward if instead of power moments we use
\emph{normalized factorial moments}. The $m$th normalized factorial moment
is defined as the expectation
\begin{equation}
  \E\condt[\Bigg]{
    \binom{n \ya}{m}\,\binom{n}{m}^{-1}}{\yH}
  \equiv
  \sum_{\ya} %m!\,
  \binom{n \ya}{m}\,  \binom{n}{m}^{-1}\, \pf(\ya \| \yH),
\end{equation}
where $\ya\in\set{0, 1/n, 2/n, \dotsc, (n-1)/n, 1}$ is the normalized total
activity of a population of $n$ neurons, and $\pf(\ya \| \yH)$ is the
probability for activity $\ya$ given assumptions $\yH$. This moment can be
interpreted as the expectation of the number of distinct $m$-tuples of
simultaneously active neurons, normalized by the maximum possible number of
distinct $m$-tuples. Note that the first $m$ factorial moments determine
the first $m$ moments and vice versa, so fixing the ones is equivalent to
fixing the others. Now assume that these $n$ neurons are a sample from a
full population of $N$ neurons, with possible normalized activity
$\yA \in\set{0,1/N,\dotsc,1}$. We have this extremely convenient property:
\emph{The first $n-1$ normalized factorial moments for the sample and for
  the full population are numerically identical}:
\begin{equation}
  \label{eq:equal_m_factorialmoments}
  \E\condt[\big]{\tbinom{n\ya}{m}\;\tbinom{n}{m}^{-1}}{\yH} =
  \E\condt[\big]{\tbinom{N\yA}{m}\;\tbinom{N}{m}^{-1}}{\yH},
  \qquad m < n.
\end{equation}

Hence we can apply the maximum-entropy method to obtain a distribution for
the full population by constraining, for any $m$s we please, its $m$th
factorial moment to be equal to the sample's recorded average of
$\tbinom{n\ya}{m}\;\tbinom{n}{m}^{-1}$. If $\ya_{t}$ is the normalized
total activity of the sample during the time bin $t$, during a recording
with $T$ time bins, the constraint on the distribution $\pf(\yA \| \yH)$
for the full population is
\begin{equation}
  \label{eq:final_constraint}
  \frac{1}{T}\sum_{t}\binom{n\ya_{t}}{m}\,\binom{n}{m}^{-1}
  =
   \sum_{\yA}
  \binom{N \yA}{m}\,  \binom{N}{m}^{-1}\, \pf(\yA \| \yH)
\end{equation}
for any $m$s we wish.
